---
title: "Homework 7"
author: "Seth Marceno"
date: "6/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
load('HW7.RData')
library(car)
```

#Question 1

Here I read in the data and do forwards AIC selection:
```{r, warning = FALSE}
data(mantel)
X1 <- mantel$X1
X2 <- mantel$X2
X3 <- mantel$X3
Y <- mantel$Y
fullModel <- ~ X1 + X2 + X3
baseModel <- lm(Y ~ X1)
forward <- step(baseModel, fullModel, direction = 'forward')
```

Here we do backwards AIC selection:
```{r, warning=FALSE}
m1 <- update(baseModel, fullModel)
backwards <- step(m1, scope = c(lower = ~ X1) , direction = 'backward')
```

Here I do forwards BIC selection:
```{r, warning = FALSE}
forwardBIC <- step(baseModel, fullModel, direction = 'forward', k = log(length(X1)))
```

Here I do backwards BIC selection:
```{r, warning=FALSE}
backwardBIC <- step(m1, scope = c(lower = ~ X1), direction = 'backward', k = log(length(X1)))
```
Based off of both forwards and backwards selection using BIC and AIC, we see that the active regressors are X1 and X2.

#Question 2

##Case 1:
```{r}
e <- 1
hii <- .9
p <- 4
n <- 54
sigma <- 4
  
  
ri_1 <- e/ (sigma*(sqrt(1-hii)))
ri_1
Di_1 <- (1/p)*(ri_1^2)*(hii/(1-hii))
Di_1
ti_1 <- ri_1*(((n-p-1)/(n-p-(ri_1^2)))^0.5)
ti_1
qt(p = 0.975, df = 49, lower.tail = TRUE)
```
At level alpha = 0.05, our critical value is 2.01. Since ti_1 is not greater than our critical value, we determine that this point is not influential. 

##Case 2:

```{r}
e <- 1.732
hii <- 0.75
p <- 4
n <- 54
sigma <- 4
  
  
ri_2 <- e/ (sigma*(sqrt(1-hii)))
ri_2
Di_2 <- (1/p)*(ri_2^2)*(hii/(1-hii))
Di_2
ti_2 <- ri_2*(((n-p-1)/(n-p-(ri_2^2)))^0.5)
ti_2
```
Comparing our test statistic to the critical value we found above, again we see our test statistic is smaller, so we conlcude that this point is not influential.

##Case 3:

```{r}
e <- 9
hii <- 0.25
p <- 4
n <- 54
sigma <- 4
  
  
ri_3 <- e/ (sigma*(sqrt(1-hii)))
ri_3
Di_3 <- (1/p)*(ri_3^2)*(hii/(1-hii))
Di_3
ti_3 <- ri_3*(((n-p-1)/(n-p-(ri_3^2)))^0.5)
ti_3
```
Since our test statistic is greater than our critical value (2.01) we determine that this point is influential. 

##Case 4

```{r}
e <- 10.295
hii <- 0.185
p <- 4
n <- 54
sigma <- 4
  
  
ri_4 <- e/ (sigma*(sqrt(1-hii)))
ri_4
Di_4 <- (1/p)*(ri_4^2)*(hii/(1-hii))
Di_4
ti_4 <- ri_4*(((n-p-1)/(n-p-(ri_4^2)))^0.5)
ti_4
```
Since our test statistic is greater than our critical value (2.01) we determine that this point is influential. 


#Question 3

##Part(a)

```{r, warning=FALSE}
data(lathe1)
Speed <- lathe1$Speed
Feed <- lathe1$Feed
Life <- lathe1$Life
fit1 <- lm(Life ~ Speed + Feed + I(Speed^2) + I(Feed^2) + Speed*Feed)
boxCox(fit1)
```
Since 0 is within the interval, we know to log transform our response. 

##Part(b)

```{r}
fit2 <- lm(log(Life) ~ Speed + Feed + I(Speed^2) + I(Feed^2) + Speed*Feed)
influenceIndexPlot(fit2)
```

Here we see points 9 and 10 have the highest cooks distance, therefore we will remove them.

```{r}
new.lathe1 <- lathe1[-c(9, 10), ]
new.Speed <- new.lathe1$Speed
new.Feed <- new.lathe1$Feed
new.Life <- new.lathe1$Life
fit3 <- lm(log(new.Life) ~ new.Speed + new.Feed + I(new.Speed^2) + I(new.Feed^2) + new.Speed*new.Feed)
summary(fit2)
summary(fit3)
```
Based off of our summary function we see that before removing the influencial points, our R squared value is 0.9702, and after we get an R squared of 0.9759. Therefore we can see the non-influential data has a slightly better fit than with the influential points. 






















```{r, include = FALSE}
save.image('HW7.RData')
```

